\documentclass[11pt,a4paper]{article}
\setcounter{secnumdepth}{6}

\usepackage{standalone}
\usepackage{geometry} % Used to adjust the document margins
\usepackage{graphicx}
\usepackage[font=footnotesize]{caption}
\usepackage[noadjust]{cite}
\usepackage[toc,page]{appendix}
\usepackage{hyperref}
\usepackage{fullpage}
\usepackage{amsmath}
\usepackage{sidecap}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{minted}
\usepackage{framed}
\usepackage[colorinlistoftodos]{todonotes}
\usepackage{csquotes}
\usepackage{parskip} % Spaces between paragraphs

\usepackage{pgffor} % foreach loops!

\usepackage[acronym]{glossaries}
% add binding margins
\geometry{twoside,bindingoffset=1cm}

\input{ipynbsupportdefinitions.tex}

\makeglossaries
\renewcommand*\abstractname{Summary}

% OUR COMMANDS
%~ 
%~ \newenvironment{pythoncode}%
%~ {\begin{framed}
%~ \begin{minted}{python}}%
%~ {\end{minted}
%~ \end{framed}}%

% http://www.tex.ac.uk/cgi-bin/texfaq2html?label=altabcr
\setcounter{MaxMatrixCols}{50}

% package name:
\newcommand{\means}{\texttt{MEANS}}
\newcommand{\pft}{\textit{p53}}
\newcommand{\py}{\texttt{python}}
\newcommand{\sympy}{\texttt{sympy}}
\newcommand{\plt}{\texttt{matplotlib}}
\newcommand{\mat}{\texttt{MATLAB}}
\newcommand{\eg}{\emph{e.g.}}
\newcommand{\ie}{\emph{i.e.}}

\newcommand{\sauliustodo}[2][]{\todo[color=cyan, #1]{\textbf{SL:} #2}}
\newcommand{\sisitodo}[2][]{\todo[color=yellow, #1]{\textbf{SF:} #2}}
\newcommand{\quentintodo}[2][]{\todo[color=red, #1]{\textbf{QG:} #2}}
\newcommand{\citationneeded}[2][]{\todo[color=brown, fancyline, #1]{\textbf{Citation Needed:} #2}}
\newcommand{\contrib}{\emph}
\begin{document}

\listoftodos
\newpage

\title{MEANS: a new python package for Moment Expansion Approximation, Inference and Simulation}
\author{Sisi Fan, Quentin Geissmann, Saulius Lukauskas.\\
\\	
\\
\\
\\
Supervised by Ann Babtie, Paul Kirk, Eszter Lakatos and Michael Stumpf\\
\\
\\
Theoretical Systems Biology Group,\\
Imperial College London
}
\date{\today}

\clearpage\maketitle
\thispagestyle{empty}
\newpage{}

\pagenumbering{roman}

\begin{abstract}
We present \means, a python package that provides stochastic modelling based on Moment Expansion Approximation (MEA),
Linear Noise Approximation and Gillespie Stochastic Simulation Algorithm in an efficient and user-friendly way.
Specifically for MEA, we have deployed scalar and three parametric moment closure methods in both univariate and multivariate forms.
We also significantly improved MEA performance, which was a limiting factor in stochastic modelling.
In addition, the package supports a collection of solvers and distance-based inference.
To enhance usability and maintainability, we implemented modular design in the package with detailed documentation,
as well as a series of interactive tutorials in IPython notebook format.
Our package also supports SBML (Systems Biology Markup Language) standard for studying available biological systems.
Apart from package development, we have conducted further analysis on MEA, and
found that higher using maximal moment order did not necessarily improve MEA accuracy.
Moreover, we discovered that MEA was more accurate in certain regions of the parameter space than others.
Some parameter sets also resulted in systematic failure of the numerical solvers.
These limitations may provide insight into improvements of simulation and inference using MEA.
We believe \means{} will a valuable package to study MEA and for researchers to model non-linear kinetic systems.
\end{abstract}

\pagebreak % move TOC to a different page from abstract
\tableofcontents

% our acronyms
\newacronym{ode}{ODE}{Ordinary Differential Equation}
\newacronym{mea}{MEA}{Moment Expansion Approximation}
\newacronym{lna}{LNA}{Linear Noise Approximation}
\newacronym{ssa}{SSA}{Gillespie Stochastic Simulation Algorithm}
\newacronym{cme}{CME}{Chemical Master Equation}
\newacronym{abc}{ABC}{Approximate Bayesian Computation}
\newacronym{sbml}{SBML}{Systems Biology Markup Language}
\newacronym{pypi}{PyPI}{the Python Package Index}

\newglossaryentry{maxord}
{
  name=max order,
  description={Maximal moment order. Max order always (regardless to the closure method) refers to the highest order of 
  moments present in an ODE system resulting from Moment Expansion Approximation. In other words, moment expansion is closed at Max order $\mathbf{+1}$
   },
  sort=max order
}
\newpage{}
\printglossaries

\newpage{}
\pagenumbering{arabic}
\input{intro}
\newpage
\input{design_considerations}
\newpage{}
\input{examples}
\newpage{}
\input{results}
\newpage{}
\input{conclus}
\newpage{}
\bibliography{report.bib}{}
\bibliographystyle{ieeetr}

\newpage{}
\begin{appendices}
\section{Documentation}
\label{sec:documentation}

Due to the large size of the documentation we decided not to attach it to the appendix of this already lengthy report.

We made the complete means documentation available to view online instead. The fully-searchable form of the documentation can be accessed at \\
\url{http://msc.bc.ic.ac.uk:2709/index.html}.

\end{appendices}
   
\end{document}


 
a